The Practical Application of Computerised Multistage Tests

### 2021-03-01 article.Rmd
Show usage of cutScore derived from cutTheta

* Each item targets one and only one latent trait
* The items are independent of one another
* The response probability increases with the ability level using the item response function desribed above

The item response function is a logistic function with a characteristic sigmoid (or 'S' shape), and can be plotted for each item. Examples of these Item Characteristic Curves (ICCs) are given here:

@@

@@

@@

To run the simulations we'll create a set of 7,650 simulated candidates with ability range -4 to 4 (stepped in increments of 0.5). Then we'll repeat for each of the six designs above, with the same item bank.

@@

#### Module A Mapping Function, $g_{A}()$

@@

```{r}
gA1 <- function(x) {
  return(gA_coeffs[6]*x^5 + gA_coeffs[5]*x^4 + gA_coeffs[4]*x^3 + gA_coeffs[3]*x^2 + gA_coeffs[2]*x + gA_coeffs[1])
}
```

So, a candidate who scores 9 out of 12 on module A would have a $\hat\theta$ = `r gA1(9)`.

@@

```{r}
df_gA <- data.frame(
  trueScore = c(seq(0,12)),
  theta = c(rep(0, 13))
)
df_gA$theta <- gA1(df_gA$trueScore)

@@

#### Scale Scoring Function, $s()$

@@

modIIF <- inner_join(iifB, iifC, by = "theta") %>%
  rename(moduleB = information.x, moduleC = information.y) %>%
  dplyr::select(theta, moduleB, moduleC) %>%
  arrange((abs(moduleB - moduleC))) %>%
  head(n = 1)
cutTheta <- modIIF$theta

ggplot(plotData, aes(x=theta, y=information)) + 
  geom_line(aes(colour=module)) +
  labs(title="Module Information Function", x="θ", y="I(θ)") +
  geom_vline(xintercept = cutTheta, linetype="dotted")
```
Here we can see that there is point of intersection between modules B and C where $\hat\theta = $ `r cutTheta`. If we read off the true score value that is equivalent for this theta estimate in Module A, then we can use this as the cut score. Below this value the candidate should be routed to Module B, above it they should be routed to Module C.

>>>>

* Each item targets one and only one latent trait
* The items are independent of one another
* The response probability increases with the ability level using the item response function described above

### Item Difficulty and the Item Characteristic Curve

For every item, the probability that the item will be answered correctly can be visualised by plotting an Item Characteristic Curves (ICCs), which uses the above logistic function.

@@

### Item Discrimination and the Item Information Function

@@

### Logits

The log odds, or *logit* of a correct response can be expressed (in the Rasch model) as $\theta - b$; so if a candidate has ability $\theta$ which happens to be equal to the difficulty of our item $b_{583}$, then the odds are even (1:1) that they will answer it correctly. The logit value here would be 0 (remember this is also the point of maximum information for the item). The logit function is the inverse of the logistic function described earlier:

$$
logit(p) = log(\frac{p}{1-p})
$$
Where $p$ is a probability.

### Test Item Lifecycle

The item response function is a logistic function with a characteristic sigmoid (or 'S' shape), and can be plotted for each item. Throughout the rest of this article we'll consider an example item which will follow a *lifecycle* from when it is first conceived, developed, rated and trialled, banked in the item bank, then delivered as part of one or more test administrations, and finally retired from use.

Of course, this item, let's call it 'Item583', will be one of many that will, over time, be added into an item bank. The item will have a number of properties and metadata that describes it. For example the name of the item writer who created it, a set of media (audio, video or image files) that is associated with it, and importantly for our purposes, the difficulty coefficient $b$. The rest of this article follows this lifecycle, in particular how it relates to the various IRT methodologies for design and validation of multistage tests.

@@

To run the simulations we'll create a set of simulated candidates with ability range -4 to 4 (stepped in increments of 0.5). Then we'll repeat for each of the six designs above, with the same item bank.

@@

#### Module A Mapping Function

@@

```{r}
gA <- function(x) {
  return(gA_coeffs[6]*x^5 + gA_coeffs[5]*x^4 + gA_coeffs[4]*x^3 + gA_coeffs[3]*x^2 + gA_coeffs[2]*x + gA_coeffs[1])
}
```

So, a candidate who scores 9 out of 12 on module A would have a $\hat\theta$ = `r gA(9)`.

@@

```{r}
df_gA <- data.frame(
  trueScore = c(seq(0,12)),
  theta = c(rep(0, 13))
)
df_gA$theta <- gA(df_gA$trueScore)

@@

#### Scale Scoring Function

@@

stage2IIFs <- inner_join(iifB, iifC, by = "theta") %>%
  rename(moduleB = information.x, moduleC = information.y) %>%
  dplyr::select(theta, moduleB, moduleC) %>%
  arrange((abs(moduleB - moduleC))) %>%
  head(n = 1)
cutTheta <- stage2IIFs$theta

ggplot(plotData, aes(x=theta, y=information)) + 
  geom_line(aes(colour=module)) +
  labs(title="Module Information Function", x="θ", y="I(θ)") +
  geom_vline(xintercept = cutTheta, linetype="dotted")

cutScore <- round(df_moduleA_TS$trueScore[which.min(abs(df_moduleA_TS$theta - cutTheta))])
```
Here we can see that there is point of intersection between modules B and C where the ability estimate is `r cutTheta`. If we read off the true score value that is equivalent for this theta estimate in Module A, then we can use this as the cut score; in this case the (rounded) true score is `r cutScore`. Below this value the candidate should be routed to Module B, above it they should be routed to Module C.
